---
title: "Day2"
author: "Kristoffer Sundström"
date: "`r Sys.Date()`"
output: html_document
---

Install and load important packages:
```{r}
library(tidyverse) # Imports and loads the package tidyverse.
```


**************************** PART 1 **************************************

First, lets set retrieve the dataset from my GitHub repo and transform it:
```{r}
data <- read.table("~/GitHub/Advent-of-Code/2024/Inputs/Day1.txt",
           header = FALSE) # Reads the table from the GitHub directory.
```


Now lets sort the two vectors into two independent, and sorted, integer 
vectors:
```{r}
x1 <- data$V1[order(data$V1)] # Vector 1.
x2 <- data$V2[order(data$V2)] # Vector 2.
```

Lets calculate and print the differences between each of these, in absolute 
values, and sum it up: 
```{r}
Answer1 <- sum(abs(x1 - x2)) # Sums the absolute values between x1 and x2.
print(Answer1) # Prints answer for part 1 of day 1.
```

**************************** PART 2 **************************************

Now lets create a new column, named "CountV1", in which each value in column  is 
multiplied by the number of times it shows up in the second column:
```{r}
data <- data %>%
  mutate(CountV1 = sapply(V1, function(x) sum(V2 == x))) # Dplyr application in which the times each value in V1 appears in V2 are counted and reported in a new column named CountV1.
```

Now lets calculate and print the answer for part 2 by multiplying the values in 
column 1 by the new count column and sum the values togheter:
```{r}
Answer2 <- data %>%
  mutate(Similarity_Score = V1 * CountV1) %>% # Create a similarity score according to the assignment by multiplying the value in V1 with CountV1.
  pull(Similarity_Score) %>% # Pulls the Similarity_Score from data.  
  sum(data$Similarity_Score) %>% # Sums the similarity score to get the answer for part 2 of day 1.
  print() # Prints answer.

```
