Install and load necessary packages:
```{r}
library(dplyr)
```

**************************** PART 1 **************************************

First, lets set retrieve the dataset from my GitHub repo and transform it:
```{r}
data <- read.table("~/GitHub/Advent-of-Code/2024/Inputs/Day1.txt",
           header = FALSE)
```


Now lets sort the two vectors into two independent, and sorted, integer 
vectors:
```{r}
x1 <- data$V1[order(data$V1)]
x2 <- data$V2[order(data$V2)]
```

Lets calculate and print the differences between each of these, in absolute 
values, and sum it up: 
```{r}
Answer1 <- sum(abs(x1 - x2))
print(Answer1)
```

**************************** PART 2 **************************************

Now lets create a new column, named "CountV1", in which each value in column  is 
multiplied by the number of times it shows up in the second column:
```{r}
data <- data %>%
  mutate(CountV1 = sapply(V1, function(x) sum(V2 == x)))
```

Now lets calculate and print the answer for part 2 by multiplying the values in 
column 1 by the new count column and sum the values togheter:
```{r}
Answer2 <- data %>%
  mutate(Similarity_Score = V1 * CountV1) %>%
  pull(Similarity_Score)%>%
  sum(data$Similarity_Score) %>%
  print()

```
