---
title: "Day2"
author: "Kristoffer Sundström"
date: "`r Sys.Date()`"
output: html_document
---

Install and load important packages:
```{r}
library(tidyverse)
```


** PART 1 **

Import the dataset:
```{r}
data <- read.table("~/GitHub/Advent-of-Code/2024/Inputs/Day2.txt", # Reads the table from the GitHub directory.
           header = FALSE, # Indicates that the data has no headers.
           sep = " ", # Data is separated by a blank space.
           fill = NA) # If no value is present, report it as NA.
```

Now lets try to check whether the row is has a positive trend or negative trend:
```{r}
for (i in 1:7) {
  # Positive trend called "diffpos":
  data[[paste0("diffpos", i)]] <- (data[[paste0("V", i+1)]] - data[[paste0("V", i)]]) <= 3 &
                                        (data[[paste0("V", i+1)]] - data[[paste0("V", i)]]) > 0

  # Negative trend called "diffneg":
  data[[paste0("diffneg", i)]] <- (data[[paste0("V", i)]] - data[[paste0("V", i+1)]]) <= 3 &
                                        (data[[paste0("V", i)]] - data[[paste0("V", i+1)]]) > 0
}
```

Now lets check if all columns for either diff_pos or diff_neg is only TRUE,
this would indicate a safe report:
```{r}
# Check if all values in columns starting with "diffpos" are TRUE for each row:
data <- data %>%
  mutate(
    AllTruePos = apply(data[, grepl("^diffpos", names(data))], 1, function(row) {
      all(row == TRUE, na.rm = TRUE)  # Checks if all values in the diffpos columns are TRUE, ignoring NA.
    }),
    
    # Check if all values in columns starting with "diffneg" are TRUE for each row
    AllTrueNeg = apply(data[, grepl("^diffneg", names(data))], 1, function(row) {
      all(row == TRUE, na.rm = TRUE)  # Checks if all values in the diffneg columns are TRUE, ignoring NA.
    })
  )

# Convert TRUE/FALSE to 1/0 if needed:
data$AllTruePos <- ifelse(data$AllTruePos, 1, 0)
data$AllTrueNeg <- ifelse(data$AllTrueNeg, 1, 0)

# Calculate the sum of AllTruePos and AllTrueNeg:
Answer1 <- sum(data$AllTruePos) + sum(data$AllTrueNeg)

Answer1
```

** Part 2 **
First lets find the cases in which there is only one faulting report:
```{r}
# Check the number of FALSE observations in columns starting with "diffpos":
data <- data %>%
  mutate(
    PosF = apply(data[, grepl("^diffpos", names(data))], 1, function(row) {
      sum(row == FALSE, na.rm = TRUE)
    }),
    
    # Check the number of FALSE observations in columns starting with "diffpos":
    NegF = apply(data[, grepl("^diffneg", names(data))], 1, function(row) {
      sum(row == FALSE, na.rm = TRUE)
    })
  )

# Convert the values to SAFEpos and SAFEneg IF AllTruePos and AllTrueNeg are TRUE or if PosF or NegF are equal to 1.
data$SAFEpos <- ifelse(data$AllTruePos == TRUE | data$PosF == 1, 1, 0)
data$SAFEneg <- ifelse(data$AllTrueNeg == TRUE | data$NegF == 1, 1, 0)
```

Now lets remove the FALSE observations from these cases and report them as NA:
```{r}
# Filter and get new dataset for Positive trends with 1 failure:
dataPos <- data %>%
  filter(SAFEpos == 1)

# Filter and get new dataset for Negative trends with 1 failure:  
dataNeg <- data %>%
  filter(SAFEneg == 1)

# Now lets transform all FALSE reports to NAs for both new datasets:
dataPos <- dataPos %>%
  mutate(across(everything(), ~ ifelse(. == FALSE, NA, .)))

dataNeg <- dataNeg %>%
  mutate(across(everything(), ~ ifelse(. == FALSE, NA, .)))
```

Now lets rerun the code from part 1 after removing the FALSE obsvervations:
```{r}
# Check if all values in columns starting with "diffpos" are TRUE for each row given new datasets:
dataPos <- dataPos %>%
  mutate(
    AllTruePos2 = apply(dataPos[, grepl("^diffpos", names(dataPos))], 1, function(row) {
      all(row == TRUE, na.rm = TRUE)  # Checks if all values in the diffpos columns are TRUE, ignoring NA.
    }))
    
# Check if all values in columns starting with "diffneg" are TRUE for each row given new datasets:
dataNeg <- dataNeg %>%
  mutate(
    AllTrueNeg2 = apply(dataNeg[, grepl("^diffneg", names(dataNeg))], 1, function(row) {
      all(row == TRUE, na.rm = TRUE)  # Checks if all values in the diffneg columns are TRUE, ignoring NA.
    }))

# Convert TRUE/FALSE to 1/0 if needed:
dataPos$AllTruePos2 <- ifelse(dataPos$AllTruePos2, 1, 0)
dataNeg$AllTrueNeg2 <- ifelse(dataNeg$AllTrueNeg2, 1, 0)

# Calculate the sum of AllTruePos and AllTrueNeg:
Answer2 <- sum(dataPos$AllTruePos2) + sum(dataNeg$AllTrueNeg2)

Answer2
```

